I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-06-14 paper_io_sim.tex

	\section{Method}
		In this section, we discuss our approach to model page cache and
		the cache eviction strategy implemented in Linux. We also detail
		the design the simulators in order to generalize page cache as well
		as the implementation in Python and SimGrid framework. Finally, we
		describe some experimental scenarios to compare the results of
		Python simulator, baseline SimGrid simulator and the results from
		the real pipelines.

@@

			When modeling the I/O time, we consider the key factors of the
			I/O mechanism having the biggest impacts on data read/write.
			The first factor taken into account is the use of the page
			cache in Linux kernel. The second factor  is the cache eviction, 
			data flushing and cache eviction mechanisms implemented in the 
			Linux kernel, as they manage and change the status of the memory 
			including cache size, amount of available memory and 
			dirty data, based on which the kernel controls I/O activities. Therefore, 
			the idea of our simulator is to simulate the memory, the Linux 
			page cache with its LRU lists, storage device and imitate the data flushing 
			and cache eviction algorithm implemented in the kernel.

			At the level of detail in our simulator, the simulation models of memory 
			and storage devices are straightforward when they can be modeled with 
			predefined capacity and bandwidth. To simulate the cache eviction mechanism 
			implemented in Linux, our simulator also keeps tracks of page cache 
			LRU lists as in the kernel. However, instead of dealing with lists pages, we 
			consider a sequence of continuously accessed pages of a file as a single 
			unit, a block of data. A data block keeps the information about file name, 
			block size, last access and a dirty flag describing whether the data is 
			dirty or not.

			The idea of data block offers our simulator some obvious advantages. 
			In cache eviction algorithm, which is one of the key components 
			in our model, maintaining and handling pages in packet level are 
			non-trivial when they require significant implementation overhead and 
			would likely to hinder the performance of the simulator. The concept of data 
			block not only simplifies our cache eviction simulation but also allows to 
			preserve the accuracy of the simulator by using a single page as the 
			representative for a sequence of pages with similar properties including 
			file name, last access, dirty or not. Because a block represents a set of 
			pages in real systems, it can be split if some data needs to be moved 
			between the page cache LRU lists or evicted from cache. However, blocks can 
			not be merged as we want to preserve this level of granularity to maintain 
			high accuracy.

			Having these simulation components, we can generalize I/O activities in
			our simulator. When the kernel reads and writes data, it simply 
			reads/writes data from/to disk or cache with predefined disk bandwidth 
			or memory bandwidth. The data blocks are added or updated in page cache 
			LRU lists accordingly. In flushing and periodical flushing, the kernel 
			traverses through the inactive and active lists to reduce the amount of 
			dirty data, while cache eviction algorithm deletes blocks from the 
			inactive list. The page cache LRU lists are updated after every I/O action.

			\begin{figure}
   				\centering
   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
   				\caption{The additional layer between user application layer and 
   				original SimGrid API}\label{fig:interaction}
			\end{figure}

			We encapsulate our simulation components with kernel-like functionalities 
			into a layer between user simulated application and storage simulation 
			component (Figure~\ref{fig:interaction}). Our new layer includes two main
			classes, the \textbf{\textit{IOController}}, which models the kernel 
			I/O controlling logic, and the \textbf{\textit{MemoryManager}}, which 
			manages page cache LRU lists and stores all necessary information about 
			memory including capacity, amount of cache, amount of memory used 
			by applications. Because every node has its own memory, storage devices 
			and system configurations, there is one IOController object and one 
			MemoryManager object created for each node. The IOController class 
			keeps system parameters (dirty{\_}ratio, dirty{\_}background{\_}ratio), 
			implements necessary I/O control functionalities such as read, write,
			flushing, and cache evicting.

			With this layer, user simulated applications send requests to an 
			IOController of the node where requested data is stored, instead of 
			sending read/write requests directly to storage model as in simulators 
			with the current SimGrid version. Based on the memory status in 
			MemoryManager, the IOController orchestrates cache read/write, 
			disk read/write, flushing and cache eviction to fulfill application 
			requests. The time for each action is estimated and the total time of 
			a request is returned to user applications.

			\begin{algorithm}\caption{Read}\label{alg:read}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{fn}{file name}
        				\Desc{fs}{file size}
						\Desc{rt}{current simulated run time}
						\Desc{mm}{MemoryManager object}
						\Desc{$r_m$}{memory read bandwidth}
						\Desc{$r_d$}{disk read bandwidth}
   					\EndInput
   					\Output
						\Desc{rt}{simulated run time after the file is read}
   					\EndOutput
					\State mem\_required = max(0, 2*fs - mm.cached\_data(fn) - 
					mm.available\_memory()) 
					\State mem\_avail = max(0, 2 * fs - mm.cached\_data(fn) - 
					mm.dirty\_data())
					\State rt = rt +  mm.flush(mem\_required - 
					mem\_avail)	
					\State free\_mem\_required = max(0, 2 * fs - mm.cached\_data(fn) - 
					mm.free\_memory())
					\State mm.evict(free\_mem\_required) 
					\If {mm.cached\_data(fn) $>$ 0}
    					\State mm.cache\_read(fn) 
						\If {mm.cached\_data(fn) == fs}
    						\State time = mm.cache\_data(fn) / $r_m$
    						\State IOController.periodic\_flush(time) 
							\State rt = rt + time
						\EndIf
					\EndIf
					\If {mm.cached\_data(fn) $<$ fs}
						\State rt = rt + IOController.flush(mm.dirty\_data())
						\State mm.disk\_read(fs - mm.cache\_data(fn), fn)
    					\State rt = rt + (fs - mm.cache\_data(fn)) / $r_d$
					\EndIf

@@

			The algorithm to generalize file read is described in 
			\textit{Algorithm~\ref{alg:read}}. In Linux systems, data can be read in 
			two ways, read from cache with memory bandwidth and read from disk 
			with disk bandwidth. Because total read time can be seen as a 
			linear function of the read time in each of these two modes, 
			we can generalize reading process as two separate and sequential phases, 
			one is cache read, and the other is disk read. When a user application 
			requests a file read, it calls the read method of the IOController. Then, 
			IOController calculates how much memory is required to read the file 
			based on the amount of available memory, file size, and the amount of 
			file data cached (line 10). If there is not enough available memory, 
			dirty data is flushed to disk to make more memory available by calling flush
			function with the amount of memory needed as the parameter (line 11-12).
			IOController then evicts old blocks from the inactive list to
			accommodate the requested file and flushing time is simulated (line 13-14). 
			After that, IOController starts estimating file read time by first checking 
			if a some file data is cached (line 15). If there is cached data, 
			it is read from cache and the blocks containing this data are updated in 
			the page cache LRU lists with the last access time (line 16). If all file data 
			is cached, time to read from cache is added to total run time, and 
			periodical flushing is triggered to flush dirty data during the cache read 
			(line 18-19). If only a part of the file is in cache, disk read is required, 
			this does not allow flushing and makes the cache read time neglectable. 
			At the beginning of a disk read, dirty data is flushed, then data is read 
			and added to the inactive list, disk read time is estimated with disk read 
			bandwidth and added to total run time (line 23-27).

			\textit{Algorithm~\ref{alg:write}} describes our
			generalization of file write. Similar to the file read process,
			a file can also be written with two different bandwidths,
			memory bandwidth and disk bandwidth, and thus file write can be
			generalized with two modes. File can be written with memory
			bandwidth before the page cache is saturated, or the amount of
			dirty data reaches dirty\_ratio. Firstly, if dirty\_ratio is not reached 
			the amount of data can be written with memory bandwidth is
			calculated with the formula:
			\begin{equation}
				M = max(0, \frac{(d*C - D)*w_m}{w_m - w_d})\label{equa:freeamt}
			\end{equation}

@@

			By this formular, we assume that data is written with with memory 
			bandwidth until all file data is written or dirty\_ratio is reached, 
			the kernel concurrently writes data to cache and flushes dirty data 
			to disk. If file size is smaller than or equal to $M$, the whole file is 
			written to cache as dirty data with memory bandwidth (line 12). Otherwise, 
			the remaining file data is written with disk bandwidth. Before simulating 
			cache write, IOController evicts the page cache to accommodate new data 
			(line 13). Then, data is written to cache as dirty data with memory 
			bandwidth write time is added to simulated time (line 14-16). Because 
			cache write and data flushing are concurrent, the actual amount of written 
			dirty data reduces by the amount of flushed data (line 15). In the 
			second phase, in which data is written with disk bandwidth, our simulator 
			can generalize not only the case that data is written directly to disk, 
			but also the case that applications have to wait for dirty data 
			being flushed. Before writing, if the amount of free memory is not enough 
			for the remaining file data, evictable memory is evicted (line 20-23). 
			Here, the kernel does not need to guarantee that the amount of free 
			memory after eviction can accommodate the remaining data. In the case 
			free memory is less than remaining data, we the kernel can flush, 
			evict and write new data concurrently. Finally, IOController simulates 
			the time to write the remaining amount data, adds written data to 
			page cache LRU lists, and the amount of dirty data remains the same 
			(line 24). The total simulated time of the second phases is added and 
			returned to user applications.

			\begin{algorithm}\caption{Write}\label{alg:write}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{fn}{file name}
        				\Desc{fs}{file size}
						\Desc{rt}{current simulated run time}
						\Desc{mm}{MemoryManager object}
						\Desc{$w_m$}{memory write bandwidth}
						\Desc{$w_d$}{disk write bandwidth}
   					\EndInput
   					\Output
						\Desc{rt}{simulated run time after the file is read}
   					\EndOutput
					\State dirty\_left = $d$ * mm.available\_memory() - mm.dirty\_data()
					\If {dirty\_left $>$ 0}
   						\State free\_amt = min(fs, $M$) \Comment{$M$ in Equation
   						~\ref{equa:freeamt}}
    					\State IOController.evict(free\_amt - mm.free\_memory())
    					\State free\_time = free\_amt / $w_m$
    					\State mm.write(fn, free\_amt, dirty = free\_amt - free\_time * $w_d$)
   						\State rt = rt + free\_time
    				\EndIf

					\State throttled\_amt = fs - free\_amt
					\State cache\_required = max(0, throttled\_amt  - mm.free\_memory())
					\State evicted\_amt = min(cache\_required, mm.evictable\_memory())
					\State mm.evict(evicted\_amt)
					\State mm.write(fn, cached = min(mm.free\_memory(), 
					throttled\_amt), dirty = 0)
					\State rt = rt + throttled\_amt /$w_d$

			Data flushing is another key function in our simulator. It takes 
			the input as the amount of flushed data and the returned value is the 
			simulated flushing time. To flush dirty data, IOController simply 
			traverses the sorted inactive and active list, reduces the dirty data 
			amount by unmarking dirty blocks until the flushed amount reached 
			the amount passed in or there is no dirty data left in cache. 
			Similarly, periodical flushing is the function that simulates data flushing 
			during CPU time and cache read. However, the flushed data amount is 
			calculated with the input which is flushing time. In both flushing 
			functions, the last block being flushed is split if its data is not 
			entirely flushed.

			Equally important, cache eviction function is called from both 
			read and write to free up the page cache when free memory 
			is not enough. The amount of evicted data is passed in as in flushing, 
			however, this function does not add up the simulated time as 
			cache eviction time is neglectable in real systems . When eviction 
			is called, IOController deletes the last accessed blocks from the 
			inactive list until the evicted amount reaches the amount passed in or 
			there is no evictable memory left. If the block is not entirely deleted, 
			the block is split.

			Last but not least, we need to modified page cache LRU lists after 
			every I/O activity. When IOController reads the data that is not cache, 
			the data is read and added to the inactive list. If the data is already 
			in cache, the cached blocks are moved to the active list with the 
			last access time updated. Finally, MemoryManager makes the page cache 
			LRU lists balanced by moving blocks from the longer list to the shorter 
			and sorting the lists.

@@

			Firstly, to validate our simulation model, we create a simple 
			simulator independent to simulation frameworks and libraries. 
			This enables us to evaluate the accuracy and correctness of our 
			model in a simple use case before integrating it in SimGrid. 
			This simulator simulates a single-threaded pipeline running on a system 
			with a single core CPU, one local storage device and all input and 
			output files are stored locally. The results from the simulator is compared 
			and validated with the results from a real pipeline running on
			a real system.

			Having our model validated, we create other different
			simulators of different use cases using the current version of
			SimGrid and the SimGrid version that is extended with our
			model. Then, we compare the results of the simulators with
			original SimGrid and extended SimGrid with the results of real
			pipelines on a real system. The results are compared in two
			different aspects, task completion time and I/O time, as well
			as the ability of the simulator in generalization of the page
			cache in terms of dirty data and cache used.

@@

			When designing experiments, we consider a range of scenarios
			that have different impacts on the results. The experiments are
			a number of pipelines of sequential tasks. Each task reads an
			input file, performs some computation, generates an output file
			and writes it to disk. The output file of the previous task is
			the input file of the next task. Thus, each pipeline has only one
			actual input file.

			Initially, we want to have a detailed view of the effects of data size 
			on I/O behaviors and validate our simulator in modeling and 
			capturing those impacts. Thus, there should be only one pipeline and 
			the size of input and output files should vary in a range that 
			the change in file size can cause significant changes on I/O behaviors.

			Furthermore, the number of pipelines should also vary. Having our 
			simulator validated on a single threaded experiment, the next goal 
		    is to evaluate our simulation model with a multi-threaded experiment, 
		    with a varying number of pipelines run on a multi-core host and write 
		    to the same local storage device. As a result, the shared bandwidth 
		    the pipelines is different with different numbers of concurrent processes. 
		    By doing so, we can evaluate the performance of our model on the simulation 
		    of multi-threaded, shared bandwidth applications.

		    Moreover, different storage types, which include local and shared 
		    storage, should also be taken into account when they are commonly used 
		    in computer systems. The scenario of the next experiment is the same as 
		    the multi-threaded experiment, but should replace local storage device 
		    with a shared file systems, so that we can not only access our 
		    simulator on I/O with shared file system, but also compare the 
		    performance of the simulator on different storage types.

		    Finally, we adopt a real neuroimaging pipeline to evaluate the 
		    applicability of our simulator in simulating I/O time in real
			applications. We simulate the pipeline with a simulated workflow running 
			on a simulated plaform and compare the simulation results with the real 
			results. Because our work focuses on I/O time, we assume
			that CPU time is correctly modeled and use the CPU time measured 
			in the real pipelines to setup our simulated workflows.

@@

				In this experiment, our goal is to validate our model with a simple 
				simulator implemented in Python and compare the simulation precision 
				of our simulation model and the existing model in SimGrid. 
				Thus, we simulate one single thread pipeline of three tasks with the
				Python simulator, SimGrid simulator and run on our real system. 
				All files are read and written locally. We use the input sizes of 
				20 GB, 50 GB, 75 GB and 100 GB to run the pipeline on the cluster, 
				measure the CPU time of the pipeline with each input to simulate 
				the pipeline with the simulators.

@@

				In the second experiment, we evaluate the simulator in concurrent I/O 
				simulation with a multi-threaded application. As we run the pipelines on 
				only one compute node of the cluster, there are upto 32 cores per node 
				and all input and output files are read and written on the same disk,  
				we create 32 input files with the size of 3GB each and vary the number of 
				concurrent pipelines from 1 to 32. Because our goal in this experiment is 
				to evaluate concurrent I/O in our simulation model, and multi-threaded 
				simulator implementation of the Python is expensive, we only compare the 
				performance of simulators with original SimGrid, SimGrid integrated with 
				our model and and real pipelines. The results are compared in total 
				makespan of the pipelines, cumulative read time and cumulative write time.

>>>>

	\section{Method}
		In this section, we present our approach to model file read/write, 
		I/O in memory, data flushing and cache eviction mechanisms 
		implemented in Linux. 
		We also detail the design the simulators in order to simulate I/O activities 
		and the implementation in Python and SimGrid framework. 
		Finally, we describe some experimental scenarios to evaluate our model 
		with a Python simulator, a baseline SimGrid simulator, a simulator 
		with SimGrid integrated with our model, and the results from real pipelines.

@@

			In our I/O simulation model, we try to simulate the write back page cache 
			mechanisms including memory I/O, cache eviction and data flushing as we 
			believe that they are the most important factors that affect the I/O time.

			Memory is modeled with capacity, read bandwidth and write 
			bandwidth. When data is read from or written to memory, the actual 
			read/write time is calculated with the amount of data and memory 
			read/write bandwidth:

			\begin{align*}
				& t_{mr} = D / r_m \\ 
				& t_{mw} = D / w_m\
			\end{align*}

			where:
			\begin{itemize}
				\item $t_{mr}$ is time to read data from memory
				\item $t_{mw}$ is time to write data to memory
				\item $D$ is the amount of data to read/write
				\item $r_m$ is memory read bandwidth
				\item $w_m$ is memory write bandwidth
			\end{itemize}

			We introduce the concept of data block as a unit to represent cached data. 
			A data block is a subset of pages of a file in page cache that is accessed 
			at a specific time. Therefore, a block has the information about file name, 
			block size, last access and a dirty flag with values of 0 and 1, 
			0 represents block of lean data, while 1 indicates that the data block 
			is dirty. 
			Cached data of a file can be represented by multiple blocks. 
			Because a data block represents a subset of cached pages, 
			it can be split into smaller blocks, and multiple blocks can also 
			be merged into a larger one.

			In Linux kernel, the page cache LRU lists are lists of pages of 
			files in cache. In our simulator, instead of using lists of pages, 
			we simulate these lists with two lists of data blocks, an active list and 
			an inactive list, ordered by block last access as is shown in 
			Figure~\ref{fig:lrulist}.
			As in the kernel, in our simulator, we keep the size of the active list not 
			larger than twice the size of the inactive list by moving least recently 
			used blocks from the active list to the inactive list.

			\begin{figure}
   				\centering
   				\includegraphics[width=\columnwidth]{figures/lru_lists.pdf}
   				\caption{Simulation of page cache LRU lists with data blocks}\label{fig:lrulist}
			\end{figure}

			Modeling page cache LRU lists as lists of blocks offers our simulator 
			some obvious advantages. 
			Maintaining and handling lists of pages in the simulator, which is in 
			packet level, are non-trivial when they require significant implementation 
			overhead and would likely to hinder the performance of the simulator. 
			Therefore, using page cache lists of data blocks can simplify the 
			simulation of cache eviction and data flushing while preserving
			the accuracy in i/o time simulation.

			With the simulation model of page cache LRU lists of blocks, now we can 
			model I/O in memory.
			At the time when a file is accessed, some data of the file can be in 
			cache or not. 
			If accessed data is already cached, there are existing data blocks 
			representing cached data in two page cache LRU lists. 
			As cached data is accessed again, these blocks are moved to 
			the active list, dirty blocks are merged into a bigger dirty block, 
			while clean data blocks are merged into another clean block. 
			If accessed data is not in cache, new blocks representing uncached data 
			are created and put into the top of the inactive list. 
			For more detail when uncached data is accessed, in the case of 
			reading file, as all uncached data is clean data, only one clean data 
			block is added to the inactive list.
			But in the case of writing uncached file data though cache, 
			this written data can be either dirty or clean. 
			Thus, at most two data blocks can be added to the inactive lists, 
			one is clean and one is dirty. 
			We also keep the size of the active list not larger than twice the size of 
			the inactive list as in the kernel. This is done by moving least recently 
			used blocks in the active list to the inactive list when the active list 
			grows. If only a part of a data block needs to be moved, 
			the block is split.

			Next, we simulate data flushing mechanism, which writes back 
			dirty data to disk. 
			The data flushing simulation function traverses the sorted inactive first, 
			and then the active list, looks for least recently used dirty blocks and 
			sets dirty flags of these blocks to 0 until the total flushed amount 
			reaches the amount to be flushed or there is no dirty data left in cache. 
			If the last flushed block is not entirely flushed, it is split into 
			two blocks, one is flushed and one remains dirty.
			The flushing time is calculated as:

			\begin{align*}
				t_f = A / w_d
			\end{align*}

			where:
			\begin{itemize}
				\item $t_f$ is the flushing time
				\item $A$ is the amount of data to be flushed
				\item $w_d$ is the disk write bandwidth
			\end{itemize}

			For periodical flushing, because dirty data can be flushed during CPU time 
			or cache read with disk write bandwidth, the amount of flushed data is 
			calculated with disk bandwidth and flushing duration.

			Similarly, our cache eviction simulation function frees up the page cache 
			by traversing and deleting least recently used clean data blocks 
			in the inactive list.
			The amount of evicted data is passed in and data blocks are deleted 
			from the inactive list until total evicted data reaches the amount 
			passed in, or there is no clean block left in the list.
			If the last evicted block is not entirely evicted, the block is split, 
			and only one block is deleted.
			Our cache eviction simulation does not add up to the simulated time 
			since cache eviction time is negligible in real systems.

			We encapsulate the models of memory and page cache LRU lists in 
			\textit{MemoryManager} class. 
			The functionalities of this class are exposed to the simulation  
			of I/O activities including file read and file write. 
			We create \textit{IOController} class which simulates file read and 
			file write by interacting with the MemoryManager class, 
			exposing APIs to user simulated applications. 
			These two class act as an extra layer with kernel-like functionalities as 
			illustrated in Figure~\ref{fig:interaction}. 
			Simulated applications can simply request for a file read 
			or file write by calling the APIs of this layer instead of 
			calling directly to storage simulation models.
			Because every computer has its own memory and system configurations, 
			one IOController object and one MemoryManager object are created 
			for each simulated machine.

			\begin{figure}
   				\centering
   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
   				\caption{The additional layer between user application layer and 
   				storage simulation models}\label{fig:interaction}
			\end{figure}

			To read or write a file, a simulated application sends a request to the 
			IOController instance on the node where the requested file is stored.
			Based on the memory status in MemoryManager, the IOController 
			orchestrates cache read/write, disk read/write, flushing and cache eviction 
			to fulfill I/O requests, calculates total simulated time and returns to 
			simulated applications.

			\begin{algorithm}\caption{File read simulation}\label{alg:read}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{fn}{file name}
        				\Desc{fs}{file size}
						\Desc{rt}{current simulated run time}
						\Desc{mm}{MemoryManager object}
						\Desc{$r_m$}{memory read bandwidth}
						\Desc{$r_d$}{disk read bandwidth}
   					\EndInput
   					\Output
						\Desc{rt}{simulated run time after the file is read}
   					\EndOutput
					\State req = max(0, 2 * fs - mm.cached(fn) - mm.avail\_mem) 
					\State avail = max(0, 2 * fs - mm.cached(fn) - mm.dirty)
					\State rt = rt + mm.flush(req - avail)	
					\State req = max(0, 2 * fs - mm.cached(fn) - mm.free\_mem)
					\State mm.evict(req) 
					\If {mm.cached(fn) $>$ 0}  \Comment{Read file cached data}
    					\State mm.cache\_read(fn) 
						\If {mm.cached(fn) == fs} \Comment{If all file data is cached}
    						\State time = mm.cached(fn) / $r_m$
    						\State mm.periodic\_flush(time) 
							\State rt = rt + time
						\EndIf
					\EndIf
					\If {mm.cached(fn) $<$ fs} \Comment{Read uncached data from disk}
						\State mm.disk\_read(fs - mm.cached(fn), fn)
    					\State rt = rt + (fs - mm.cached(fn)) / $r_d$
					\EndIf

					\State \Return rt

			The algorithm used to simulate file read is described in 
			Algorithm~\ref{alg:read}. 
			In Linux systems with write back storage devices, data can be either read 
			from cache with memory bandwidth or read from disk with disk bandwidth. 
			Therefore, total read time can be considered as a linear function 
			of the read time with each of these bandwidths, 
			enabling us to generalize reading process as two separate phases, 
			cache read and disk read. 
			In our simulator, when a user simulated application reads a file, 
			it sends a file read request to IOController. 
			Then, IOController calculates the amount of required memory to read 
			the file based on the amount of available memory, file size, and 
			the amount of cached data of that file (line 10). 
			If there is not enough available memory, dirty data is flushed to disk 
			to increase the amount of memory available by calling flush function 
			of MemoryManager with the amount of flushed data, flushing time is 
			simulated and added to total run time (line 11-12).
			After data is flushed, MemoryManager is asked to evict data 
			from cache to accommodate the requested file (line 13-14). 
			Next, if file is cached either partially or entirely, MemoryManager is 
			called to re-access and update this cached data (line 16). 
			If all file data is cached, periodical flushing is called is to flush 
			dirty data during the cache read, and time to read from cache is added to 
			total run time, (line 18-20). 
			If file is not cached or partially cached, a disk read is required (line 23) 
			This blocks flushing and makes the cache read time neglectable. 
			Finally, data is read from disk to cache and managed 
			by MemoryManager, disk read time is estimated with disk read 
			bandwidth and added to total run time (line 24-25).

			\begin{algorithm}\caption{File write simulation}\label{alg:write}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{fn}{file name}
        				\Desc{fs}{file size}
						\Desc{rt}{current simulated run time}
						\Desc{mm}{MemoryManager object}
						\Desc{$w_m$}{memory write bandwidth}
						\Desc{$w_d$}{disk write bandwidth}
   					\EndInput
   					\Output
						\Desc{rt}{simulated run time after the file is read}
   					\EndOutput
					\State dirty\_left = $d$ * mm.avail\_mem - mm.dirty
					\If {dirty\_left $>$ 0}
   						\State free\_amt = min(fs, $M$) \Comment{$M$ in Equation
   						~\ref{equa:freeamt}}
    					\State mm.evict(free\_amt - mm.free\_mem)
    					\State free\_time = free\_amt / $w_m$
    					\State mm.write(fn, free\_amt, dirty = free\_amt - free\_time * $w_d$)
   						\State rt = rt + free\_time
    				\EndIf

					\State throttled\_amt = fs - free\_amt \Comment{Written with disk bandwidth}
					\State req = max(0, throttled\_amt  - mm.free\_mem)
					\State evicted\_amt = min(req, mm.evictable)
					\State mm.evict(evicted\_amt)
					\State cached\_amt = min(mm.free\_mem, throttled\_amt)
					\State mm.write(fn, cached = cached\_amt, dirty = 0)
					\State rt = rt + throttled\_amt /$w_d$

@@

			Algorithm~\ref{alg:write} describes our simulation of file write. 
			Similar to the file read simulation, a file can also be written with 
			two different bandwidths, memory bandwidth and disk bandwidth. 
			Files can be written with memory bandwidth before the amount 
			of dirty data reaches dirty\_ratio. Thus, our algorithm initially checks 
			this condition (line 10), and estimates the maximum amount 
			of data can be written with memory bandwidth with the following 
			formula:

			\begin{equation}
				M = max(0, \frac{(d*C - D)*w_m}{w_m - w_d})\label{equa:freeamt}
			\end{equation}

@@

			In this formula, we assume that data is written with with memory 
			bandwidth until dirty\_ratio is reached, the kernel writes data 
			to cache and flushes dirty data to disk concurrently. 
			If file size is smaller than or equal to $M$, the whole file is 
			written to cache as dirty data with memory bandwidth (line 12). 
			Otherwise, the remaining file data is written with disk bandwidth (line 20). 
			Before simulating cache write, MemoryManager is asked to evict 
			the page cache to accommodate written data (line 13). 
			Then, data is written to cache as dirty data with memory 
			bandwidth, and write time is added to total simulated time (line 14-16). 
			Because cache write and data flushing are concurrent in our algorithm, 
			the actual amount of written dirty data reduces by the amount 
			of data during cache write time (line 15). 
			If not all data is written with memory write bandwidth, the remaining data 
			is written with disk write bandwidth. This can simulate both cases that 
			data is written directly to disk, or applications wait for dirty data 
			to be flushed, and evicted to write new data. 
			If the amount of free memory is not enough for the remaining file data, 
			evictable memory is evicted (line 20-23). 
			Here, the simulator does not guarantee that the amount of free memory 
			after eviction can accommodate the remaining data. 
			In the case that free memory is less than remaining data, the amount to be 
			cached is limited by the amount of free memory (line 24). 
			However, the kernel still flushes, evicts and writes new data concurrently, 
			and remaining data is still written to cache with disk write bandwidth.
			Thus, we can simulate the time to write the remaining data with 
			disk bandwidth and ask MemoryManager to add clean cached data (line 25). 
			The simulated time to write width disk bandwidth is added to total 
			simulated run time and returned to simulated applications (line 26).

@@

			To validate our simulation model, we create a simple 
			simulator which is independent to simulation frameworks and libraries. 
			This enables us to evaluate the accuracy and correctness of our 
			model in a simple use case before integrating it to SimGrid. 
			This simulator simulates a single-threaded pipeline running on a system 
			with a single core CPU, one local storage device and all input and 
			output files are stored locally. The results from the simulator is compared 
			and validated with the results from a real pipeline running on
			a real system.

			Having our model validated, we create simulators for different use cases 
			using the current version of SimGrid and the SimGrid version that is 
			extended with our model. We leverage Wrench framework, which is 
			built on top of SimGrid to simulate our pipelines and system.
			Then, we compare the results of the simulators with
			original SimGrid and extended SimGrid with the results of real
			pipelines on a real system.

@@

			When designing experiments, our goal is to evaluate our simulation model 
			in single-threaded and multi-threaded applications which access data on 
			on different file system types. 
			To do that, we create a pipeline of sequential tasks, each task 
			reads an input file, increases every byte of the input, and writes the 
			output to disk. The output of the previous task is the input of 
			the next task. We also adopt an use case in real applications and 
			to evaluate the applicability of our simulation model.

			The simplest scenario is a single-threaded application, where memory 
			bandwidth and disk bandwidth are not shared with another thread. 
			This allows us to compare the detailed results in terms of read/write time, 
			memory states after every single task and ,more importantly, validate our 
			simulation model.

			For multi-threaded applications, we do the experiment with a varying 
			number of concurrent pipelines running on a multi-core host 
			and writing to the same local storage device. 
		    By doing so, we can assess the performance of our model in the case  
		    of concurrent I/O with shared bandwidth on a local storage device.

		    Since different storage types should also be taken into account, 
		    we conduct the next experiment with the same scenario as 
		    in the multi-threaded experiment, except that files are stored on a 
		    shared file system instead of local disk. 
		    In this way, we can not only evaluate our simulation model on I/O 
		    with shared file system, but also compare the 
		    performance of the simulator on different storage types.

		    Finally, we adopt a real pipeline in a real neuroimaging applications. 
		    We simulate the pipeline with a simulated workflow in SimGrid 
			and compare the simulation results with the real results. 
			Because our work focuses on I/O time, we assume that CPU time is 
			correctly modeled and use the CPU time measured in the real pipelines 
			to setup our simulated workflows.

@@

				In this experiment, our goal is to validate our model by comparing 
				the simulation results of the simple Python simulator with results of 
				a simulator implemented with original SimGrid, a simulator 
				implemented with SimGrid integrated with our model, and a real pipeline 
				running of the cluster. 
				The pipeline consists of three tasks, each task reads a file, 
				increases every byte and writes output file to local disk. 
				We use the input sizes of 20 GB, 50 GB, 75 GB and 100 GB, run 
				the pipeline on the cluster, measure the CPU time of the pipeline 
				with each input to simulate the pipeline with the simulators.

@@

				In the second experiment, we evaluate the simulator in concurrent I/O 
				with a multi-threaded application. As we run the pipelines on 
				only one compute node of the cluster, which has upto 32 cores per node,  
				we create 32 input files with the size of 3GB each and vary the number of 
				concurrent pipelines from 1 to 32. 
				Because multi-threaded simulator implementation in Python is 
				expensive, and we have our model integrated in SimGrid, 
				we only compare the results of simulators with original SimGrid, 
				SimGrid integrated with our model and and real pipelines. 
				The results are compared in total makespan of the pipelines, 
				cumulative read time and cumulative write time.
